---
title: "Week 6 Discussion Assignment"
author: "Lucas Hennessey"
date: "10/3/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## To get ready to investigate breast cancer data we need to load our libraries and import the data

```{r loading libraries and data}

# opening our libraries, maybe more than I need

library(data.table)
library(dplyr)
library(car)
library(caret)
library(caTools)
library(pROC)
library(MASS)
library(tidyverse)

# Importing our dataset and converting to a data frame.  Used notepad to add column names before importing 

breast.cancer.wisconsin <- read.csv("~/Python_class/Week 6 Logistic Regression Stuff/breast-cancer-wisconsin.data")

dt <- as.data.table(breast.cancer.wisconsin)

# setting the seed for later analysis

set.seed(1)

```

## I'm going to take a quick look at our data, then clean and prepare it for our regresssion

```{r EDA}

# removing id

dt <- dt[, c("id") := NULL]

#checking structure of df

str(dt)

# checking summary statistics

summary(dt)

# converting data types to allow for logistic regression

dt$bare_nucleoli <- as.integer(dt$bare_nucleoli)

dt$diagnosis <- factor(dt$diagnosis, labels =c('benign', 'malignant'))

str(dt)

# removing NAs

dt <- dt[complete.cases(dt)]

# finally split our data into a training and a test set

samp <- sample.split(dt$diagnosis, SplitRatio = 0.8)
train <- subset(dt, samp == TRUE)
test <- subset(dt, samp == FALSE)

```

## Now to model our cleaned, prepped data on our training set and take a look


```{r modeling}

# creating the model

model <- glm(diagnosis ~ ., data = train, family = "binomial")

# looking at the results

summary(model)

# checking for collinearity (assumption checking)

vif(model)

# performing step AIC to remove high p values

stepAIC(model, dirrection = 'both')

# se cell shape, cell shape and cell size don't appear to be helpful in our regression, so I'll remove them and model again

model <- glm(formula = diagnosis ~ clump_thickness + marginal_adhesion + bare_nucleoli + bland_chromatin + normal_nucleoli + mitoses, family = "binomial", data = train)

summary(model)

# predicting on our training data using our simplified model

trainpreds <- predict(model, type = 'response', train)

# setting prediction values.  Assuming that '2' means negative and '4' means positive

trainp <- factor(trainpreds >= 0.5, labels = c('benign', 'malignant'))

# Building a confusion matrix to evaluate our model

trainCM <- confusionMatrix(train$diagnosis, trainp)
trainCM

```
## So far our simplified model looks fantastic.  Let's test it against our test data set

```{r predictions}

# predict on the test data            
testpreds <- predict(model, type = 'response', test)

# set cutoff labels for predictions
testp <- factor(testpreds >= 0.5, labels = c('benign', 'malignant'))

# Build a confusion matrix to see results
testCM <- confusionMatrix(test$diagnosis, testp)
testCM

```

## Our model looks really good against the test data too, so we'll create ROC curves for our model on the training set and our model on the test set and we'll look at confusion matrices for all of our models to evaluate the performance of each

```{r roc curves}

# ROC curve for the Train data
train_roc_curve <- roc(train$diagnosis, trainpreds)
train_roc_curve
plot(train_roc_curve)
train_rocc <- coords(roc=train_roc_curve, x = 'best', best.method = 'closest.topleft')
train_rocc

# ROC curve for the Test data
test_roc_curve <- roc(test$diagnosis, testpreds)
test_roc_curve
plot(test_roc_curve)
test_rocc <- coords(roc=test_roc_curve, x = 'best', best.method = 'closest.topleft')
test_rocc

# predict on training data using ROC cut-off
trainrocp <- factor(trainpreds >= as.numeric(train_rocc[1]), labels = c('benign', 'malignant'))

# confusion matrix
trainROCCM <- confusionMatrix(train$diagnosis, trainrocp)
trainROCCM

# Predict on the test data            
testp <- factor(testpreds >= as.numeric(test_rocc[1]), labels = c('benign', 'malignant'))

# confusion matrix
testROCCM <- confusionMatrix(test$diagnosis, testp)
testROCCM

#View all the Confusion matrices
trainCM
trainROCCM

testCM
testROCCM


```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
